The yaq Project: \\ Standardized Software Enabling Flexible Instrumentation

### 2022-11-03 paper.tex
hyphenate self-describing

In \yaq{}, communication between daemons and clients is performed over TCP/IP using Apache Avro RPC \cite{AvroSpecification}.
Avro provides an agreed upon standard for efficient serialization of data and method calls from a remote (client) process.
Practically, the \yaq{} interface looks like a collection of methods or functions, which Avro calls ``messages''.
Each message has defined input parameters and output return types.
A sensor might implement a message called ``\texttt{get\_measured}'' which takes no parameters and returns a dictionary mapping channel names to numeric or array data.
A motor would implement a pair of messages for setting and reading back the motor position: ``\texttt{set\_position(float position)} $\rightarrow$ \texttt{null}'' and ``\texttt{get\_position()} $\rightarrow$ \texttt{float}''.
These self describing messages make up the lowest level functionality of the \yaq{} interface.
Each individual communication between client and daemon involves one message being requested by the client and the response returned from the daemon.
Each daemon supports a collection of messages for its unique functionality, called the ``protocol''.

@@

In the middle, the same three hardware peripherals are represented in an interactive, graphical form.
\texttt{yaqc-qtpy}\cite{yaqc-qtpy} is a graphical application which builds interactive controls based on traits for any conceivable \yaq{} protocol.
The self describing \yaq{} interface is used to provide graphical elements for the most commonly used messages.
This is an invaluable tool which provides a ``free'' graphical user interface (GUI) to any daemon.

@@

We have created several tools to aide in daemon development.
First, a Python library, \texttt{yaqd-core}\cite{yaqd-core-python}, which implements shared functionality.
Second, \texttt{yaq-traits}\cite{yaq-traits} is a command line application which allows the description of messages provided by a \yaq{} protocol to be written in a human-readable fashion and translated into a more fully described machine readable format.
The format it generates is an important part of how \yaq{} protocols are self describing.
This shields developers from the details of Apache Avro, which can be somewhat esoteric.

>>>>

In \yaq{}, communication between daemons and clients is performed over TCP/IP using Apache Avro RPC \cite{AvroSpecification}.
Avro provides an agreed upon standard for efficient serialization of data and method calls from a remote (client) process.
Practically, the \yaq{} interface looks like a collection of methods or functions, which Avro calls ``messages''.
Each message has defined input parameters and output return types.
A sensor might implement a message called ``\texttt{get\_measured}'' which takes no parameters and returns a dictionary mapping channel names to numeric or array data.
A motor would implement a pair of messages for setting and reading back the motor position: ``\texttt{set\_position(float position)} $\rightarrow$ \texttt{null}'' and ``\texttt{get\_position()} $\rightarrow$ \texttt{float}''.
These self-describing messages make up the lowest level functionality of the \yaq{} interface.
Each individual communication between client and daemon involves one message being requested by the client and the response returned from the daemon.
Each daemon supports a collection of messages for its unique functionality, called the ``protocol''.

@@

In the middle, the same three hardware peripherals are represented in an interactive, graphical form.
\texttt{yaqc-qtpy}\cite{yaqc-qtpy} is a graphical application which builds interactive controls based on traits for any conceivable \yaq{} protocol.
The self-describing \yaq{} interface is used to provide graphical elements for the most commonly used messages.
This is an invaluable tool which provides a ``free'' graphical user interface (GUI) to any daemon.

@@

We have created several tools to aide in daemon development.
First, a Python library, \texttt{yaqd-core}\cite{yaqd-core-python}, which implements shared functionality.
Second, \texttt{yaq-traits}\cite{yaq-traits} is a command line application which allows the description of messages provided by a \yaq{} protocol to be written in a human-readable fashion and translated into a more fully described machine readable format.
The format it generates is an important part of how \yaq{} protocols are self-describing.
This shields developers from the details of Apache Avro, which can be somewhat esoteric.
